---
title: 'Chapter 5: Data Transformation'
author: "Doron Dusheiko"
date: "10 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```
## Exercises 5.2.4

> Find all flights that had an arrival delay of two or more hours

```{r}
flights
```

```{r}
flights %>%
  filter(arr_delay >= 120)
```

> Flew to Houston (IAH or HOU)

```{r}
flights %>%
  filter(dest %in% c("IAH", "HOU"))
```

> Were operated by United, American, or Delta

Method 1: 

```{r}
 codes <- airlines %>% 
  filter(grepl("United|American|Delta", name)) %>%
  select(carrier)

flights %>% 
  filter(carrier %in% codes$carrier)

```

Method 2: using joins

```{r}
flights %>%
  inner_join(airlines, by = "carrier") %>%
  filter(grepl("United|American|Delta", name))
```

> Departed in summer (July, August, and September)

```{r}
flights %>%
  filter(month %in% c(7, 8, 9))
```

> Arrived more than two hours late, but didn’t leave late

```{r}
flights %>%
  filter(arr_delay > 120, dep_delay <= 0)
```

> Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
flights %>%
  filter(dep_delay >= 60, dep_delay - arr_delay > 30)
```

> Departed between midnight and 6am (inclusive)

```{r}
flights %>%
  filter(dep_time <= 600 | dep_time == 2400)
```

> Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

```{r}
flights %>%
  filter(between(month, 7, 9))
```

> How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
sum(is.na(flights$dep_time))
```

```{r}
count_missing <- function(col) {
  sum(is.na(col))
}

flights %>%
  map_int(count_missing) %>%
  sort.int(decreasing = TRUE)

```

> Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

```{r}
NA ^ 0
NA | TRUE
FALSE & NA
NA * 0
```

