---
title: 'Chapter 5: Data Transformation'
author: "Doron Dusheiko"
date: "10 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```
## Exercises 5.2.4

> Find all flights that had an arrival delay of two or more hours

```{r}
flights
```

```{r}
flights %>%
  filter(arr_delay >= 120)
```

> Flew to Houston (IAH or HOU)

```{r}
flights %>%
  filter(dest %in% c("IAH", "HOU"))
```

> Were operated by United, American, or Delta

Method 1: 

```{r}
 codes <- airlines %>% 
  filter(grepl("United|American|Delta", name)) %>%
  select(carrier)

flights %>% 
  filter(carrier %in% codes$carrier)

```

Method 2: using joins

```{r}
flights %>%
  inner_join(airlines, by = "carrier") %>%
  filter(grepl("United|American|Delta", name))
```

> Departed in summer (July, August, and September)

```{r}
flights %>%
  filter(month %in% c(7, 8, 9))
```

> Arrived more than two hours late, but didn’t leave late

```{r}
flights %>%
  filter(arr_delay > 120, dep_delay <= 0)
```

> Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
flights %>%
  filter(dep_delay >= 60, dep_delay - arr_delay > 30)
```

> Departed between midnight and 6am (inclusive)

```{r}
flights %>%
  filter(dep_time <= 600 | dep_time == 2400)
```

> Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

```{r}
flights %>%
  filter(between(month, 7, 9))
```

> How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
sum(is.na(flights$dep_time))
```

```{r}

flights %>%
  map_int(~ sum(is.na(.))) %>%
  sort.int(decreasing = TRUE)

```

> Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

```{r}
NA ^ 0
NA | TRUE
FALSE & NA
NA * 0
```

## Exercises 5.3.1

> How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()`).

```{r}
arrange(flights, desc(is.na(dep_time)), dep_time)
```

> Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r}
flights %>%
  arrange(desc(dep_delay))

flights %>%
  arrange(dep_delay)
```

> Sort flights to find the fastest flights.

```{r}
flights %>%
  arrange(air_time) %>%
  select(air_time, everything())
```

> Which flights travelled the longest? Which travelled the shortest?

```{r}
flights %>%
  arrange(desc(distance))

flights %>%
  arrange(distance)
```

## Exercises 5.4.1

> Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

```{r}
flights %>%
  select(starts_with("dep"), starts_with("arr"))

flights %>%
  select(ends_with("time"), ends_with("delay"), -starts_with("sched"), -air_time)

flights %>%
  select(dep_time:arr_delay, -starts_with("sched"))
```

> What happens if you include the name of a variable multiple times in a select() call?

```{r}
flights %>%
  select(dep_time, dep_time)
```

> What does the one_of() function do? Why might it be helpful in conjunction with this vector?

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")

flights %>%
  select(one_of(vars))
```

> Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```

```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```

## Exercises 5.5.2

> Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
flights %>%
  mutate(
    dep_time = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
     sched_dep_time = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% 1440
  )
```

  
> Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
flights %>%
  mutate(total_time = arr_time - dep_time) %>%
  select(air_time, total_time)
```

Need to convert these fields to minutes as well

```{r}
flights %>%
  mutate(
    arr_time = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440,
    dep_time = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    total_time = arr_time - dep_time
  ) %>%
  select(air_time, total_time)
```

> Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

Would expect that `dep_time = sched_dep_time + dep_delay`. 

```{r}
flight_mins <- flights %>%
  mutate(
    sched_dep_time = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% 1440,
    dep_time = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    dep_time_diff = dep_time - sched_dep_time - dep_delay
  ) %>%
  select(dep_time, sched_dep_time, dep_delay, dep_time_diff)

```

```{r}
filter(flight_mins, dep_time_diff != 0)
```

The time difference of 1440 mins suggestes that those flights took off close to midnight so dep_time > arr_time. Let's check they are the only possible values. 

```{r}
unique(filter(flight_mins, dep_time_diff != 0)$dep_time_diff)
```

> Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r}
flights %>%
  mutate(rank = min_rank(desc(dep_delay))) %>%
  filter(rank <= 10) %>%
  arrange(rank) %>%
  select(rank, origin, dest, dep_delay, dep_time:dest, -c(flight, tailnum))
```

> What does 1:3 + 1:10 return? Why?

```{r}
1:3
1:10
1:3 + 1:10
```

After computing matching pairs, the seq with less values is recycled acrcoss the sequence with more values.

> What trigonometric functions does R provide?

see `?Trig`
