---
title: 'Chapter 3: Data Visualisation'
author: "Doron Dusheiko"
date: "10 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Exercises 3.2.4

> Question 1: Run `ggplot(data = mpg)`. What do you see?

```{r}
ggplot(data = mpg)
```

Answer: This gives us the plotting surface without any coordinate system or data points.

> Question 2: How many rows are in mpg? How many columns?

```{r}
glimpse(mpg)
```

Answer: There are 234 rows and 11 columns of data. 

> Question 3: What does the drv variable describe? Read the help for ?mpg to find out.

`drv` describes the drive train of the vehicle, which can have categorical values, f = front-wheel drive, r = rear wheel drive, 4 = 4wd

> Question 4: Make a scatterplot of hwy vs cyl

```{r}
ggplot(mpg, aes(cyl, hwy)) + 
  geom_point()
```

As the number of cylinders increases, the fuel efficiency decreases. 

> Question 5: What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r}
ggplot(mpg, aes(class, drv)) +
  geom_point()
```

Both are categorical variables which makes it hard to read this plot for insight. 

## Exercises 3.3.1

> Question 1: What’s gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

The current code is trying to group the data by a non-existant variable "blue" rather than changing the colour of the points to blue. The following code resolves the issue.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

> Queestion 2: Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

```{r}
mpg
```

Categorical: manufacturer, model, trans, drv, fl, class
Continuous: displ, cty and hwy. year and cyl are numerical but should probably be convereted to categorical unless one wants to perform numerical computations based on them. 

> Question 3: Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

 
```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy, colour = year))

ggplot(mpg) + 
  geom_point(aes(displ, hwy, size = year))

# ggplot(mpg) + 
#   geom_point(aes(displ, hwy, shape = year))
```

Mapping colour to a continous variable translates to a range over which the observations vary. Size creates an increase in size of the points and shape is not possible. This makes sense since there are an infinite number of divisions for a continuous variable so how would you know which shape to map. 

> Question 4: What happens if you map the same variable to multiple aesthetics?

```{r}
ggplot(mpg) + 
  geom_point(aes(drv, cty, color = drv))

ggplot(mpg) + 
  geom_point(aes(displ, hwy, color = displ))
```

This allows us to more clearly see the varying relationship at particular combination of values.

> Question 5: What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use `?geom_point`)

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = cty), shape = 21, size = 5, fill = "red", stroke = 2)
```
`stroke` lets us put a border around the plotted point. The value we provide indicates the thickness of the border. It should be specified outside of `aes()` otherwise it will attempt to map the shape to a specific variable range.

> Question 6: What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy, color = displ < 5))
```

This lets us apply conditional filtering to the colour bracketing. In this case, if the displacement of the engine is less than 5 show in one colour, if it's equal to or greater than 5 show in another colour. This is a great way to quickly identify patterns in the data.

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy, color = displ < 3 | displ > 5, shape = displ < 3 | displ > 5))
```

## Exercises 3.5.1

> Question 1: What happens if you facet on a continuous variable?

```{r}
ggplot(mpg, aes(class, cty)) +
  geom_point() +
  facet_wrap(~ displ)
```

You'll get a new subplot for each increment of the continuous variable which doesn't help you much. 

> Question 2: What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

They represent combinations which have no values, such as 5 cylinder, rear-wheel drive car. In the latter plot, this is visible by the lack of a point at that combination.

> Question 3: What plots does the following code make? What does . do?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

First plots drv on the rows, second plots cyl on the cols.

> Question 4: Take the first faceted plot in this section: What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

It can be easier to spot patterns if you break subplots according to a certain categorical variable. If it's all on one plot seperated by colours and there are many levels, then it may be hard to distinguish patterns. If the dataset is large then faceting is probably better unless there are too many levels and you end up with tens of subplots. If the dataset is small and the number of levels is small, a single plot seperated via a colour aesthetic may be better.

> Question 5: Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have `nrow` and `ncol` arguments?

nrow and ncol allow you to specify how many subplots occur in each row and column respectively. This allows you to tweak the visual layout of the overall plot.

> Question 6: When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

As the number of levels and subplots grows, it becomes easier to read off the plots horizontally rather than vertically. 

## Exercises 3.6.1

> Question 1: What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

geom_line, geom_boxplot, geom_histogram, geom_area

> Question 2: Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
Prediction: a scatter plot showing relationship between highway efficiency and engine size, seperated by the drive train of the vehicle. This will be overlaid with a fitted curve that doesn't show the standard error.
Actual: Both the points and the fitted curve are individually coloured. 

> Question 3: What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

It prevents a legend from being plotted. Removing it will then make use of the default value which is show.legend = TRUE, so a legend will be shown. It was used earlier so that the graph wasn't confused by adding a legend at the end, this wouldn't have made sense for the other two plots.

> Question 4: What does the se argument to geom_smooth() do?

Displays a confidence interval or standard error around the curve.

> Question 5: Will these two graphs look different? Why/why not?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```
Prediction: They will look the same, both referrring to the same underlying data and variables, the second has more duplication.

> Question 6: Recreate the R code necessary to generate the following graphs.

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(mpg, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE, aes(group = drv))

ggplot(mpg, aes(displ, hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes( color = drv)) + 
  geom_smooth(se = FALSE)

ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = drv)) + 
  geom_smooth(se = FALSE, aes(linetype = drv))

ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = drv)) + 
  geom_point(shape = "circle open", color = "white", stroke = 3) 
```

## Exercises 3.7.1

> What is the default geom associated with `stat_summary()`? How could you rewrite the previous plot to use that geom function instead of the stat function?

geom_histogram/geom_freqpoly

```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )

ggplot(data = diamonds) +
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat = "summary",
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )

```

> What does geom_col() do? How is it different to geom_bar()?

geom_col is used when the values in the data represent the actual heights of the bar. geom_bar is better when the heights of the bar are to be computed from the data.

> Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

TODO

> What variables does stat_smooth() compute? What parameters control its behaviour?

It computes a linear regession model between the variables, an attempt to fit the the data to a smooth curve. It's controlled by its method of computation, the number of points to evaluate for the smoother and various others.

> In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```

If group is not set to 1, then all the bars have prop == 1. The function geom_bar() assumes that the groups are equal to the x values, since the stat computes the counts within the group.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop.., group = color))
```

## Exercises 3.8.1

> What is the problem with this plot? How could you improve it?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

There is overlap on many points, we can add jitter to improve it.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

> What parameters to geom_jitter() control the amount of jittering?

`width` and `height` control the horizontal and vertical jitter respectively. Th jitter is added in both directions, i.e. left and right for width and top and bottom for height. 

> Compare and contrast geom_jitter() with geom_count()

geom_count counts overlapping points. It gives you a sense of how much overplotting occurs at a specific point.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()
```


> What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

Default positions is `dodge2` which preserves the vertical position of an geom while adjusting the horizontal position.

## Exercises 3.9.1

> Turn a stacked bar chart into a pie chart using coord_polar()

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill") +
  coord_polar(theta = "y")
```

